---
title: "Instacart Exploratory Analysis"
author: "Laia Esquerrà"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}
products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?

```{r}
dbGetQuery(
  sc,
  "
SELECT order_dow AS DayOfWeek, COUNT(1) AS Frequency 
FROM orders_tbl
GROUP BY order_dow
ORDER BY Frequency DESC
LIMIT 1
  "
  )
```

* The most popular hour of doing order?

```{r}
dbGetQuery(
  sc,
  "
SELECT order_hour_of_day AS Hour, COUNT(order_hour_of_day) AS Frequency 
FROM orders_tbl
GROUP BY order_hour_of_day
ORDER BY Frequency DESC
LIMIT 1
  "
  )
```


* The most reordered products?

```{r}
dbGetQuery(
  sc,
  "
SELECT a.product_id, product_name, SUM(reordered) AS n_reorders 
FROM order_products__prior_tbl AS a
  LEFT JOIN products_tbl AS b ON a.product_id=b.product_id
GROUP BY a.product_id, b.product_name
ORDER BY n_reorders DESC
  "
  )
```

<<<<<<< HEAD
* The less reorderd products?
=======
* The less reorderd product?
>>>>>>> 9ea9d20648ebb0b776f393b61880b8be9a963f8c

```{r}
dbGetQuery(
  sc,
  "
<<<<<<< HEAD
SELECT a.product_id, product_name, SUM(reordered) AS n_reorders 
FROM order_products__prior_tbl AS 
  LEFT JOIN products_tbl AS b ON a.product_id=b.product_id
GROUP BY a.product_id, b.product_name
HAVING n_reorders > 0
ORDER BY n_reorders ASC)
"
)
=======
SELECT *
FROM (SELECT a.product_id, product_name, SUM(reordered) AS n_reorders 
      FROM order_products__prior_tbl AS a
        LEFT JOIN products_tbl AS b ON a.product_id=b.product_id
      GROUP BY a.product_id, b.product_name
      ORDER BY n_reorders ASC)
WHERE n_reorders != 0
  "
  )
>>>>>>> 9ea9d20648ebb0b776f393b61880b8be9a963f8c
```

* Plot histogram of number of products in an order.

```{r}
products_order_txt <- "
SELECT n_prod
,   COUNT(order_id) AS n_orders
FROM (
  SELECT order_id, COUNT(1) AS n_prod
  FROM order_products__prior_tbl
  GROUP BY order_id
) a
GROUP BY n_prod
"

products_order <-
  dbGetQuery(sc, products_order_txt)

products_order
```

```{r}
products_order %>%
  filter(n_prod <= 50) %>%
  ggplot(
    aes(n_prod, n_orders)) +
  geom_bar(stat="identity", fill='turquoise') +
  scale_y_continuous(label=scales::comma) +
  xlab("Order") +
  ylab("Number of products") +
  theme_minimal()
```

```{r}
order_products__prior %>%
  group_by(order_id) %>%
  summarise(n_prod = n()) %>%
  group_by(n_prod) %>%
  summarise(n_orders = n())%>%
  filter(n_prod <= 50) %>%
  collect %>%   #li hem de dir que executi fins aquí per a poder continuar amb l'histograma
  ggplot(aes(n_prod, n_orders)) +
    geom_bar(stat="identity", fill='turquoise') +
    scale_y_continuous(label=scales::comma) +
    xlab("Order") +
    ylab("Number of products") +
    theme_minimal()
```

* What is the most popular department? 

```{r}
dbGetQuery(
  sc,
  "
SELECT department_id, COUNT(1) AS Frequency 
FROM products_tbl
GROUP BY department_id
ORDER BY Frequency DESC
LIMIT 1
"
  )
```


## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```

```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).

```{r}
dpt_prod_popularity_txt <- "
SELECT a.product_id, product_name, COUNT(1) AS popularity
FROM order_products__prior_tbl AS a
  LEFT JOIN products_tbl AS b ON a.product_id=b.product_id
WHERE department_id={{dptId}}
GROUP BY a.product_id, b.product_name
ORDER BY popularity DESC
LIMIT 10
"

<<<<<<< HEAD
data <- list(dptId=11)
=======
data <- list(dptId=19)
>>>>>>> 9ea9d20648ebb0b776f393b61880b8be9a963f8c

dpt_prod_popularity_txt %>%
  whisker.render(data) %>%
  cat
```

```{r}
dpt_prod_popularity_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .)
```

```{r}
dpt_prod_popularity_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .)%>%
  ggplot(
    aes(reorder(product_name, popularity, function(x) x), 
        popularity)) +
  geom_bar(stat="identity", fill='yellowgreen') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("Product") +
  ylab(paste("Popularity in Department", data$dptId)) +
  theme_minimal()
```

* For given day, rank department popularity.

```{r}
dpt_popularity_txt <- "
SELECT department_id, COUNT(department_id) as popularity
FROM (
  SELECT product_id
  FROM order_products__train_tbl AS a INNER JOIN
       orders_tbl AS b ON a.order_id=b.order_id
  WHERE order_dow = {{dowId}}
  ) AS ab INNER JOIN
  products_tbl AS c ON ab.product_id=c.product_id
GROUP BY department_id
ORDER BY popularity DESC
"

data <- list(dowId=0)

dpt_popularity_txt %>%
  whisker.render(data) %>%
  cat
```

```{r}
dpt_popularity_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .)
```

```{r}
dpt_popularity_txt %>%
whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(
    aes(reorder(department_id, popularity, function(x) x), 
        popularity)) +
  geom_bar(stat="identity", fill='pink') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("Department") +
  ylab(paste("Popularity on Day", data$dowId)) +
  theme_minimal()
```


## More on orders (Window Functions)

```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```

```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```

```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```

*Product that is usually orders in numerous baskets*

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

Calcula el nombre de productes que hi ha en una cistella de la compra i ho afegeix a cada producte; llavors per cada product_id calcula el nombre mig de productes que hi ha en les cistelles en què és comprat.

*Could modify the query and add product name?*

```{r}
"
SELECT b.product_name, a.*
FROM (
  SELECT pn.product_id
  ,   AVG(products_number) AS avg_products_number
  FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
  ) pn
  GROUP BY product_id
  ORDER BY avg_products_number DESC
  LIMIT 100
) a LEFT JOIN products_tbl b ON a.product_id=b.product_id
" %>%
dbGetQuery(sc, .)
```

*Product that is usually orders in small baskets*

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number ASC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

- With name

```{r}
"
SELECT b.product_name, a.*
FROM (
  SELECT pn.product_id
  ,   AVG(products_number) AS avg_products_number
  FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
  ) pn
  GROUP BY product_id
  ORDER BY avg_products_number ASC
  LIMIT 100
) a LEFT JOIN products_tbl b ON a.product_id=b.product_id
" %>%
dbGetQuery(sc, .)
```

*For the 10 first users and all their orders find the days since first order*

- Days from first to last:

```{r}
dbGetQuery(sc, 
"
SELECT user_id, SUM(days_since_prior_order) as days_since_first_order
FROM(
  SELECT user_id
  ,   order_id
  ,   days_since_prior_order
  ,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
  FROM orders_tbl
  WHERE user_id <= 10)
GROUP BY user_id
ORDER BY user_id
")
```

- Cummulative sum until each order:

```{r}
dbGetQuery(sc, 
"
SELECT user_id
,   order_id
,   order_number
,   days_since_prior_order
,   IF(days_since_first_order IS NULL, 0, days_since_first_order) as days_since_first_order
FROM(
SELECT user_id
,   order_id
,   order_number
,   days_since_prior_order
,   SUM(days_since_prior_order) OVER (partition by user_id ORDER BY order_number) AS days_since_first_order
FROM orders_tbl
WHERE user_id <= 10 
) a
ORDER BY user_id
")
```

```{r}
"
SELECT user_id
,   order_id
,   order_number
,   days_since_prior_order
,   COALESCE(
      SUM(days_since_prior_order) OVER
      (PARTITION by user_id ORDER BY order_number),
      0) AS days_since_first_order
FROM orders_tbl
WHERE user_id <= 10 
ORDER BY user_id
,   order_number
" %>%
  dbGetQuery(sql_con,.)
```


```{r}
orders %>%
  filter(user_id <= 10) %>%
  select(user_id, order_number, days_since_prior_order) %>%
  arrange(user_id, order_number) %>%
  mutate(days_since_first_order = cumsum(days_since_prior_order)) %>%
  mutate(days_since_first_order=ifelse(is.null(days_since_prior_order), 0, days_since_first_order))
```