---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
spark_memory = "3G"
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

dbGetQuery(sc, products_popularity_txt)
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

dbGetQuery(sc, products_popularity_txt)
```

```{r}
dbGetQuery(sc, products_popularity_txt) %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

dbGetQuery(sc, products_popularity_txt)
```

## Questions

* Which day of week people order most fequently?
```{r}
result <- dbGetQuery(sc, "SELECT order_dow, count(1) as num FROM orders_tbl group by order_dow order by order_dow")
ggplot(result) + geom_bar(aes(x=order_dow, y=num), stat = "identity")
print(result)
```

* The most popular hour of doing order? 
```{r}
result <- dbGetQuery(sc, "SELECT order_hour_of_day, count(1) as num FROM orders_tbl group by order_hour_of_day order by order_hour_of_day")
ggplot(result) + geom_bar(aes(x=order_hour_of_day, y=num), stat = "identity")
print(result)
```

* The most reordered products?
```{r}
result <- dbGetQuery(sc, "SELECT product_id, sum(reordered) as num FROM order_products__prior_tbl group by product_id order by sum(reordered) desc limit 100")
print(result)
```

* The less reorderd product?
```{r}
result <- dbGetQuery(sc, "select * from (SELECT product_id, sum(reordered) as num FROM order_products__prior_tbl group by product_id order by sum(reordered)) t where num>1 order by num asc limit 100")
result %>% filter(num>1) %>% head
```

* Plot histogram of number of products in an order.
```{r}
result <- dbGetQuery(sc, "SELECT order_id, count(1) as num FROM order_products__prior_tbl group by order_id")
ggplot(result) + geom_histogram(aes(x = num))
```

* What is the most popular department? 
```{r}
# result <- dbGetQuery(sc, "SELECT order_hour_of_day, count(1) as num FROM orders_tbl group by order_hour_of_day order by order_hour_of_day")
result
```


## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).
* For given day, rank department popularity.

## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```

```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```

```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```


```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```


```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT
      t1.product_id,
      t2.products_number
    FROM order_products__train_tbl t1,
     (
    SELECT order_id
    ,   COUNT(1) AS products_number
    FROM order_products__train_tbl
  GROUP BY order_id) t2
where 
t2.order_id = t1.order_id
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```


```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
,sum(case when products_number > 10 then 1 else 0 end)/count(1),
count(1)
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```



```{r}
Ex1 <- dbGetQuery(sc,"SELECT order_dow, count(1) number_of_orders
                  FROM orders_tbl 
                  GROUP BY order_dow 
                  ORDER BY count(1) desc") #d0
Ex1 %>% ggplot()+geom_col( aes(order_dow, number_of_orders) )


```


```{r}
Ex2 <- dbGetQuery(sc,"SELECT order_hour_of_day, count(1) number_of_orders 
                  FROM orders_tbl 
                  GROUP BY order_hour_of_day 
                  ORDER BY count(1) desc") #10h
Ex2 %>% ggplot()+geom_col( aes(order_hour_of_day, number_of_orders) )

```


```{r}
Ex3 <- dbGetQuery(sc,"SELECT product_id, count(1) number_of_reorders 
                  FROM order_products__prior_tbl 
                  WHERE reordered=1 
                  GROUP BY product_id 
                  ORDER BY count(1) desc")
Ex3 %>% head() %>% mutate(product_id=as.factor(product_id)) %>% ggplot()+geom_col( aes(product_id, number_of_reorders) )
Ex3 %>% tail() %>% mutate(product_id=as.factor(product_id)) %>% ggplot()+geom_col( aes(product_id, number_of_reorders) )

```


```{r}
EX5 <- dbGetQuery(sc,"SELECT order_id, count(1) number_of_products FROM order_products__prior_tbl GROUP BY order_id")

EX5 %>% ggplot()+geom_histogram(aes(number_of_products),binwidth = 1)


```


```{r}
EX6 <- dbGetQuery(sc,"SELECT department_id, count(1) number_of_products FROM order_products__prior_tbl t1 ,products_tbl t2 where t1.product_id=t2.product_id GROUP BY department_id")
EX6 %>% mutate(department_id=as.factor(department_id)) %>% ggplot()+geom_col(aes(department_id,number_of_products))







```


## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).
* For given day, rank department popularity.


## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 



```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```

```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```
* Product that is usually orders in numerous baskets

```{r}











"
SELECT product_name
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn,
 products_tbl p
where pn.product_id = p.product_id
GROUP BY product_name
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* Could modify the query and add prduct name?

* Prduct that is usualy orders in small baskets.
```{r}
"
SELECT product_name
,   AVG(products_number) AS avg_products_number
FROM (





    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number

    FROM order_products__train_tbl
) pn,
 products_tbl p
where pn.product_id = p.product_id




GROUP BY product_name
ORDER BY avg_products_number 
LIMIT 100
" %>%
dbGetQuery(sc, .)
```


*for 10 first users and all their orders find days since first order.
```{r}

"
SELECT user_id, order_number,days_since_prior_order,
SUM(days_since_prior_order) OVER(partition by user_id ORDER BY user_id,order_number ROWS BETWEEN UNBOUNDED PRECEDING AND CURRENT ROW) AS days_since_first_order
FROM orders_tbl
where user_id <=10
ORDER BY user_id" %>%
dbGetQuery(sc, .)


"
SELECT user_id, order_number,days_since_prior_order,
SUM(days_since_prior_order) OVER(partition by user_id ORDER BY user_id,order_number ROWS BETWEEN 2  PRECEDING AND CURRENT ROW) AS days_since_first_order
FROM orders_tbl
where user_id <=10
ORDER BY user_id" %>%
dbGetQuery(sc, .)



"
SELECT user_id, order_number,days_since_prior_order,
avg(days_since_prior_order) OVER(partition by user_id ORDER BY user_id,order_number ROWS BETWEEN 2  PRECEDING AND 1 PRECEDING ) AS days_since_first_order
FROM orders_tbl
where user_id <=10
ORDER BY user_id" %>%














dbGetQuery(sc, .)

```