---
title: "JESSICA PESANTEZ- Big Data Management with R"
output:
  pdf_document: default
  html_document: default
---


```{r echo=FALSE, comment=NA}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```



### EXERCISE 1

*Parejas de productos que más se compran juntos en una orden de compra. 
```{r}
agrup_prod_order<-  "
  
SELECT p.product_id as Product_id1, op.product_id as Product_id2,
COUNT(*) as Purchase_Orders 
FROM order_products__train_tbl op
INNER JOIN order_products__train_tbl p
ON op.order_id = p.order_id AND op.product_id > p.product_id
WHERE op.order_id >0
GROUP BY p.product_id, op.product_id
ORDER BY COUNT (*) DESC
"
agrup_prod_order <-
  dbGetQuery(sc, agrup_prod_order)
agrup_prod_order
```

### EXERCISE 2

* Which are the 20 most popular aisles at 12H00? And plot them. 

```{r}
aisle_popularity <- "
SELECT aisle_id, COUNT(aisle_id) AS popularidad
FROM (
  SELECT product_id
  FROM order_products__prior_tbl AS a INNER JOIN
      orders_tbl AS b ON a.order_id=b.order_id
  WHERE order_hour_of_day=12) AS ab INNER JOIN
  products_tbl AS c
  ON ab.product_id = c.product_id
  GROUP BY aisle_id
  ORDER BY popularidad DESC
LIMIT 20
"

aisle_popularity <-
  dbGetQuery(sc, aisle_popularity)
aisle_popularity
```

```{r}
aisle_popularity %>%
  ggplot(
    aes(reorder(aisle_id, popularidad, function(x) x), 
        popularidad)) +
  geom_bar(stat="identity", fill='red') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("Clasifación de las islas") +
  ylab("Popularidad a las 12H00") +
  theme_minimal()
```


### EXERCISE 3

* For a given day of the week (Wednesday), find the most popular hours of purchasing 

```{r}
hours_popularity<-dbGetQuery(sc,"SELECT order_hour_of_day , COUNT(1) AS n_Times
    FROM orders_tbl
    WHERE order_dow =3
    GROUP BY order_hour_of_day
    ORDER BY n_Times DESC
   ")
hours_popularity
```

```{r}
hours_popularity %>%
  ggplot(
    aes(reorder(order_hour_of_day, n_Times, function(x) x), 
        n_Times)) +
  geom_point(stat="identity", fill='darkred', color="darkgreen") +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("Order hour of Tuesday") +
  ylab("Frequency of Purchases") +
  theme_minimal()
```

### EXERCISE 4

* TOTAL SUM OF ORDERS THAT WERE REORDERED 

```{r}
"
SELECT 
COUNT(DISTINCT product_id) AS n_productos
FROM order_products__train_tbl
WHERE reordered =1
" %>%
dbGetQuery(sc, .)
```


#### EXERCISE 5

* Conteo del número de usuarios con respecto a los días de la semana (en este caso solo el fin de semana)

```{r}
wk_us_all<-dbGetQuery(sc,"SELECT order_dow
    ,   user_id
    ,   COUNT(user_id) OVER(partition by order_dow) AS num_users
    FROM orders_tbl
WHERE order_dow>=5
    ORDER BY num_users DESC
LIMIT 100
    ")
wk_us_all
```


```{r}
wk_us_all %>%
  ggplot(
    aes(reorder(num_users,order_dow, function(x) x), 
        order_dow)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("Days of the week") +
  ylab("Number of users") +
  theme_minimal()
```





