---
title: "Final Project"
author: "Marta Herrera"
date: "June 30, 2017"
output:
  pdf_document: default
  html_document: default
---

# Analysis of instacart

Read the data:

```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```


1. Which are the most populars products (top 10)?

```{r}
dbGetQuery(sc, "
SELECT t1.product_id, product_name, COUNT(*) AS n_orders
FROM order_products__prior_tbl t1 LEFT JOIN products_tbl t2 ON t1.product_id=t2.product_id
GROUP BY t1.product_id, product_name
ORDER BY n_orders DESC
LIMIT 10 ") %>%
  ggplot(aes(reorder(product_name, n_orders, function(x) x), n_orders)) +
  geom_bar(stat="identity", fill='darkblue') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("Products") +
  ylab("Number of orders")
```


2. How may products are never bought again?

```{r}
dbGetQuery(sc, "
SELECT count(*) as n_prod
FROM(
  SELECT product_id, sum(reordered) as n_reord
  FROM order_products__prior_tbl
  GROUP BY product_id)
WHERE n_reord=0")
```

There are 4372 products that are never bought again.


3. For each of the 5 customers that come the higher number of times, which is the probability to come again each hour of the day?

There are a lot of users with 100 orders, I'll take the first 5.

```{r}
dbGetQuery(sc, "
SELECT user_id, order_hour_of_day, n_ord/sum(n_ord) OVER(partition by user_id) as prob
FROM (
SELECT user_id, order_hour_of_day, count(*) AS n_ord
FROM orders_tbl
WHERE user_id in (
  SELECT user_id
  FROM orders_tbl
  GROUP BY user_id
  ORDER BY count(*) desc
  LIMIT 5)
GROUP BY user_id, order_hour_of_day
ORDER BY user_id, order_hour_of_day
)")
```



# Recommender system

```{r}
order_products__prior %>%
  select(order_id, product_id) %>%
  left_join(orders, by="order_id") %>%
  filter(user_id <= 10) %>% 
  select(product_id, user_id) %>%
  group_by(user_id, product_id) %>%
  summarise(rating = n()) %>%
  rename(user = user_id) %>%
  mutate(item=product_id) %>%
  select(user, item, rating) ->
  user_item_rating
```

```{r}
explicit_model <- ml_als_factorization( user_item_rating, iter.max = 5, regularization.parameter = 0.01)
```

```{r}
#matriu amb files=productes
V <- as.matrix(explicit_model$item.factors)[,-1]

#matriu amb files=individus
U <- as.matrix(explicit_model$user.factors)[,-1]

#multipliquem matrius
A <- U %*% t(V)
```


For the 10 first users, which product would we recommend them?

```{r}
index <- apply(A, 1, which.max)
V2 <- as.data.frame(explicit_model$item.factors)

recommended_prod <- c()
for (i in index){
  recommended_prod <- c(recommended_prod, V2$id[i])
}

df <- data.frame(users=1:10, recommended_prod)

df %>%
  left_join(products, by=c("recommended_prod" = "product_id"), copy=T) %>%
  select(users, recommended_prod, product_name)

```


For the user 1, which 5 products that he has never bought would we recommend him?

```{r}
#Products that he has bought
bought <- orders %>%
  filter(user_id==1) %>%
  inner_join(order_products__prior, by="order_id") %>%
  distinct(product_id) %>% 
  collect()

#remove this products from the matrix
bought <- as.vector(bought$product_id)
cols_to_remove <- c()
for (i in bought){
  cols_to_remove <- c(cols_to_remove, which(V2$id==i))
}
df_user1 <- data.frame(id_prod= V2$id[-cols_to_remove],  A=A[1,-cols_to_remove])

#Product with maximum rating
max <- df_user1[which.max(df_user1$A),] #id=32792

top5 <- df_user1 %>%
  arrange(desc(A)) %>%
  top_n(5)

products %>%
  filter(product_id %in% top5$id_prod)
```

