---
title: "Final Project"
author: "Begona Dobon"
date: "July 4, 2017"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, fig.width=5, fig.height=5, 
                      fig.pos = 'h',message=FALSE)

options("scipen"=100, "digits"=2)
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)
library(whisker)
library(magrittr)

src_tbls(sc)

```

# Is the order of the products in a basket dependent of reordering? 

## What is the average number of products in an order?

```{r}
number_products_order <-
  dbGetQuery(sc, "SELECT order_id, 
                  COUNT(1) AS n_products
                  FROM order_products__prior_tbl
                  GROUP BY order_id")

number_products_order %>%
  ggplot(aes(n_products)) + geom_histogram(binwidth = 5) 


```

An order has on average `r mean(number_products_order$n_products)` products. 

## What proportion of products in a basket are reordered? 

```{r}
proportion_reordered_products <-
  dbGetQuery(sc, "SELECT order_id, 
                  SUM(reordered) AS totalReordered,
                  MAX(add_to_cart_order) AS sizeBasket,
                  SUM(reordered)/ MAX(add_to_cart_order) AS PropReordered
                  FROM order_products__prior_tbl
                  GROUP BY order_id")

proportion_reordered_products %>%
  ggplot(aes(PropReordered)) + geom_histogram() +  
    scale_y_continuous(label=scales::comma)

```

On average, `r mean(proportion_reordered_products$PropReordered) * 100` % of the products of a basket have been bougth previously. 

An order has a median size of `r median(proportion_reordered_products$sizeBasket)`. I will use that value as a cutoff of the first items added to an order. 

## How many of the first `r median(proportion_reordered_products$sizeBasket)` products in a basket are reordered? 

```{r}
proportion_FirstReordered_txt <-"SELECT order_id, 
                  SUM(reordered) AS totalReordered,
                  MAX(add_to_cart_order) AS sizeBasket,
                  SUM(reordered)/ MAX(add_to_cart_order) AS PropReordered
                  FROM order_products__prior_tbl
                  WHERE add_to_cart_order <= {{cutoff}}
                  GROUP BY order_id"


data <- list(cutoff = median(proportion_reordered_products$sizeBasket))

proportion_FirstReordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(PropReordered)) + geom_histogram() +  
    scale_y_continuous(label=scales::comma)

```


## Do reordered products tend to be added first in the baskets?

```{r}
position_reordered_products <-
  dbGetQuery(sc, "SELECT add_to_cart_order, reordered
                  FROM order_products__prior_tbl")

position_reordered_products %>%
  ggplot(aes(factor(reordered), add_to_cart_order)) + geom_boxplot() 

```

Reordered products seem to be added first in the orders.  



# Time between orders and size of the basket

## How many days happened between orders? 

```{r}
days_between_orders <-
  dbGetQuery(sc, "SELECT user_id, 
                COUNT(1) AS n_orders,
                AVG(days_since_prior_order) AS time_between_orders
                FROM orders_tbl
                GROUP BY user_id
                ORDER BY user_id DESC")

days_between_orders %>%
  ggplot(aes(time_between_orders)) + geom_histogram() +  
    scale_y_continuous(label=scales::comma)


```

On average, a user makes an order every `r ceiling(mean(days_between_orders$time_between_orders))` days.

## How many days happened between orders? Taking into account only users that have used the app at least twice.  

```{r}
days_between_orders %>% 
    filter(n_orders >=2) %>% 
    ggplot(aes(time_between_orders)) + geom_histogram() +  
    scale_y_continuous(label=scales::comma)

```


## Do small baskets have less time between reorders? 

```{r}
size_basket_over_time_txt <-
"
SELECT user_id, op.order_id, order_number,n_products
FROM (SELECT user_id, order_id, order_number
FROM orders_tbl) op
LEFT JOIN (
SELECT order_id,
COUNT(order_id) AS n_products
FROM order_products__prior_tbl
GROUP BY order_id ) p
ON op.order_id = p.order_id
"
size_basket_over_time <- dbGetQuery(sc,size_basket_over_time_txt)

ggplot(size_basket_over_time, aes(x=order_number, y=n_products), group=user_id, col=user_id)+geom_line(color="grey")

head(proportion_reordered_products)


days_between_orders <-
  dbGetQuery(sc, "SELECT user_id, 
                COUNT(1) AS n_orders,
                AVG(days_since_prior_order) AS time_between_orders
                FROM orders_tbl
                GROUP BY user_id
                ORDER BY user_id DESC")

days_between_orders %>%
  ggplot(aes(time_between_orders)) + geom_histogram() +  
    scale_y_continuous(label=scales::comma)

```




## Are the most reordered products added first to the order?






## Do users tend to buy more, less or the same over time? Check this one!
```{r}
size_basket_over_time_txt <-
"
SELECT user_id, op.order_id, order_number,n_products
FROM (SELECT user_id, order_id, order_number
FROM orders_tbl) op
LEFT JOIN (
SELECT order_id,
COUNT(order_id) AS n_products
FROM order_products__prior_tbl
GROUP BY order_id ) p
ON op.order_id = p.order_id
"
size_basket_over_time <- dbGetQuery(sc,size_basket_over_time_txt)
head(size_basket_over_time)

ggplot(size_basket_over_time, aes(x=order_number, y=n_products), group=user_id, col=user_id)+geom_line(color="grey")


```

#####################################################

## What is the most reordered product by user? 

```{r}
most_common_product_user_txt <-
"
SELECT user_id, op.order_id,product_id
FROM (
    SELECT user_id, order_id
    FROM orders_tbl) op
LEFT JOIN (
    SELECT order_id, product_id
    FROM order_products__prior_tbl
    ) p
ON op.order_id = p.order_id

"
most_common_product_user <- dbGetQuery(sc,most_common_product_user_txt)
head(most_common_product_user)
```



```{r}

dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")

dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")


products_popularity_txt <- "SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity

```
