---
title: "Football Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---

```{r}
source("readDataToMemory.R")
readFootball()

library(ggplot2)
library(ggthemes)

<<<<<<< HEAD
##print(dbListTables(con))
```

```
select ~ SELECT
filter ~ WHERE
arrange ~ ORDER
summarise ~ aggregators: sum, min, sd, etc.
mutate ~ operators: +, *, log, etc.
```
=======
print(dbListTables(sql_con))
```

>>>>>>> upstream/master
## Pipes

```{r}
head(players)

players %>%
  head
```

```{r}
players %>%
  colnames
```


```{r}
sapply(players, class)

players %>%
  sapply(class)

# the . means giving class as a second argument
class %>%
  sapply(players, . )  
```


## SELECT ~ select

```{r}
players %>%
  head %>%
  select(player_name, birthday)
```

<<<<<<< HEAD
=======
```{r}
dbGetQuery(sql_con, "
SELECT player_name
,   birthday
FROM Player
LIMIT 6
")
```

>>>>>>> upstream/master
## WHERE ~ filter

```{r}
players %>%
  filter(birthday >= "1998") %>%
  head
```

```{r}
dbGetQuery(sql_con, "
SELECT player_name
,   birthday
FROM Player
WHERE birthday >= 1998
LIMIT 6
")
```

## ORDER ~ arrange

```{r}
players %>%
  arrange(birthday) %>%
  head
```

```{r}
dbGetQuery(sql_con, "
SELECT player_name
,   birthday
FROM Player
ORDER by birthday ASC
LIMIT 6
")
```

## aggregators ~ summarise

```{r}
players %>%
  summarise(weight=mean(weight), height=mean(height))
```

```{r}
dbGetQuery(sql_con, "
SELECT AVG(weight) AS mean_weight, 
      AVG(height) AS mean_height
FROM Player
")
```

## mutate

```{r}
players %>%
  mutate(weight_kg = weight*0.45359237) %>%
  summarise(weight=mean(weight_kg), height=mean(height))
```

```{r}
dbGetQuery(sql_con, "
SELECT AVG(weight*0.45359237) AS mean_weight, 
      AVG(height) AS mean_height
FROM Player
")
```

## GROUP BY

```{r}
players %>%
  mutate(year=as.numeric(substring(birthday, 1, 4))) %>%
  group_by(year) %>%
  summarise(weight=mean(weight)) %>%
  ggplot(aes(year, weight)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```


```{r}
dbGetQuery(sql_con, "
SELECT CAST(SUBSTR(birthday, 1, 4) AS INT) AS year, 
AVG(weight) AS mean_weight
FROM Player
GROUP BY year
") %>%
  ggplot(aes(year, mean_weight)) +
  geom_line(colour="darkorange") +
  theme_minimal()

## this is in sqlite, in hive we need to repeat in GROUP the whole expression 

# dbGetQuery(sql_con, "
# SELECT CAST(SUBSTR(birthday, 1, 4) AS INT) AS year, 
# AVG(weight) AS mean_weight
# FROM Player
# GROUP BY CAST(SUBSTR(birthday, 1, 4) AS INT)
# ") 

```




```{r}
"
SELECT CAST(SUBSTR(birthday, 1, 4) AS INT) AS year
,   COUNT(1) AS n
FROM Player
GROUP BY CAST(SUBSTR(birthday, 1, 4) AS INT) 
" %>%
dbGetQuery(sql_con, .) %>%
  ggplot(aes(year, n)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```

```{r}
dbGetQuery(sql_con, "
SELECT CAST(SUBSTR(birthday, 1, 4) AS INT) AS year, 
      COUNT(1) AS nsample
FROM Player
GROUP BY year
") %>%
  ggplot(aes(year, nsample)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```

```{r}
players %>%
  mutate(year=as.numeric(substring(birthday, 1, 4))) %>%
  group_by(year) %>%
  summarise(height=mean(height), nsamples=n()) %>%
  filter(nsamples >= 200) %>%
  ggplot(aes(year, height)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```

```{r}
### this doesn't work
dbGetQuery(sql_con, "
SELECT AVG(height) AS mean_height,
      CAST(SUBSTR(birthday, 1, 4) AS INT) AS year, 
      COUNT(height) AS nsample
FROM Player
WHERE nsample >= 200
GROUP BY year

") %>%
  ggplot(aes(year, mean_height)) +
  geom_line(colour="darkorange") +
  theme_minimal()


```
<<<<<<< HEAD
=======



Exercise

* Translate from SQL to dply or from dply to SQL
```{r}
dbGetQuery(sql_con, "
SELECT *
FROM Match
LIMIT 6
")
```

```{r}
dbGetQuery(sql_con, "
SELECT country_id
,   COUNT(1) AS n
FROM Match
GROUP BY country_id
ORDER BY n DESC
")
```
```{r}
countries
```


```{r}
games %>%
  left_join(countries, by=c("country_id" = "id")) %>%
  select(name, home_team_api_id) %>%
  group_by(name) %>%
  summarise(teams = n_distinct(home_team_api_id)) %>%
  arrange(desc(teams))
```

```{r}
dbGetQuery(sql_con, "
SELECT *
FROM Player_Attributes
LIMIT 6
")

```

* From where do players with higest accuracy come from?

* Add team name and translate to dplyr

```{r}
dbGetQuery(sql_con, "
SELECT stage, home_team_api_id, home_team_goal
FROM Match
WHERE country_id = 1
AND season = '2008/2009'
ORDER BY stage
")
```
<<<<<<< HEAD
>>>>>>> upstream/master
=======


## Countries

```{r}
countries
```

```{r}
players %>%
  head
```

>>>>>>> upstream/master
