---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)

```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?
* The most popular hour of doing order? 
* The most reordered products?
* The less reordered product?
* Plot histogram of number of products in an order.
* What is the most popular department? 

```{r}
# * Which day of week people order most fequently?

Ex1 <- dbGetQuery(sc,"SELECT order_dow, count(1) number_of_orders
                  FROM orders_tbl 
                  GROUP BY order_dow 
                  ORDER BY count(1) desc") #d0
Ex1 %>% ggplot()+geom_col( aes(order_dow, number_of_orders) )
```
```{r}
# * The most popular hour of doing order? 

Ex2 <- dbGetQuery(sc,"SELECT order_hour_of_day, count(1) number_of_orders 
                  FROM orders_tbl 
                  GROUP BY order_hour_of_day 
                  ORDER BY count(1) desc") #10h
Ex2 %>% ggplot()+geom_col( aes(order_hour_of_day, number_of_orders) )
```

```{r}
# * The most reordered products?
# * The less reordered product?

Ex3 <- dbGetQuery(sc,"SELECT product_id, count(1) number_of_reorders 
                  FROM order_products__prior_tbl 
                  WHERE reordered=1 
                  GROUP BY product_id 
                  ORDER BY count(1) desc")
Ex3 %>% head() %>% mutate(product_id=as.factor(product_id)) %>% ggplot()+geom_col( aes(product_id, number_of_reorders) )
Ex3 %>% tail() %>% mutate(product_id=as.factor(product_id)) %>% ggplot()+geom_col( aes(product_id, number_of_reorders) )
```

```{r}
# * Plot histogram of number of products in an order.
EX5a <- dbGetQuery(sc,"
SELECT order_id, count(1) number_of_products 
  FROM order_products__prior_tbl 
GROUP BY order_id") %>% 
  ggplot()+geom_bar(aes(number_of_products),binwidth = 1)+ylab("number_of_orders")



EX5b <- " 
SELECT count(1) number_of_orders,number_of_products
FROM (
      SELECT order_id, count(1) number_of_products 
        FROM order_products__prior_tbl 
    GROUP BY order_id) T 
GROUP BY number_of_products" %>% 
  dbGetQuery(sc, .)  %>% 
  ggplot()+
  geom_col(aes(number_of_products,number_of_orders))
geom_bar
EX5a
EX5b
```
```{r}
EX6 <- dbGetQuery(sc,"SELECT department_id, count(1) number_of_products FROM order_products__prior_tbl t1 ,products_tbl t2 where t1.product_id=t2.product_id GROUP BY department_id")
EX6 %>% mutate(department_id=as.factor(department_id)) %>% ggplot()+geom_col(aes(department_id,number_of_products))

```

## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).
* For given day, rank department popularity.


## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 



```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```
* Product that is usually orders in numerous baskets

```{r}
"
SELECT product_name
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn,
 products_tbl p
where pn.product_id = p.product_id
GROUP BY product_name
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* Could modify the query and add prduct name?

* Prduct that is usualy orders in small baskets.
```{r}
"
SELECT product_name
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn,
 products_tbl p
where pn.product_id = p.product_id
GROUP BY product_name
ORDER BY avg_products_number 
LIMIT 100
" %>%
dbGetQuery(sc, .)
```


*for 10 first users and all their orders find days since first order.
```{r}

"
SELECT user_id, order_number,days_since_prior_order,
coalesce(
SUM(days_since_prior_order) OVER(partition by user_id ORDER BY user_id,order_number ROWS BETWEEN UNBOUNDED PRECEDING AND CURRENT ROW)
,0)  AS days_since_first_order
FROM orders_tbl
where user_id <=10
ORDER BY user_id" %>%
dbGetQuery(sc, .)

"
SELECT user_id, order_number,days_since_prior_order,
coalesce(
    SUM(days_since_prior_order) OVER(partition by user_id ORDER BY user_id,order_number )
    ,0)  AS days_since_first_order
FROM orders_tbl
where user_id <=10
ORDER BY user_id" %>%
dbGetQuery(sc, .)


# "
# SELECT user_id, order_number,days_since_prior_order,
# SUM(days_since_prior_order) OVER(partition by user_id ORDER BY user_id,order_number ROWS BETWEEN 2  PRECEDING AND CURRENT ROW) AS days_since_first_order
# FROM orders_tbl
# where user_id <=10
# ORDER BY user_id" %>%
# dbGetQuery(sc, .)

# "
# SELECT user_id, order_number,days_since_prior_order,
# avg(days_since_prior_order) OVER(partition by user_id ORDER BY user_id,order_number ROWS BETWEEN 2  PRECEDING AND 1 PRECEDING ) AS days_since_first_order
# FROM orders_tbl
# where user_id <=10
# ORDER BY user_id" %>%
# dbGetQuery(sc, .)

```
```{r}

orders %>%
  filter(user_id <= 10) %>%
  select(user_id, order_number, days_since_prior_order) %>%
  arrange(user_id, order_number) %>%
  group_by(user_id) %>% 
  mutate(days_since_first_order = cumsum(days_since_prior_order)) %>%
  mutate(days_since_first_order=ifelse(is.null(days_since_prior_order), 0, days_since_first_order)) %>% 
  arrange(user_id, order_number) 
```

